---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# bulletsamplr
 [![license](https://img.shields.io/badge/license-MIT + file LICENSE-lightgrey.svg)](https://choosealicense.com/)
 [![CRAN status](https://www.r-pkg.org/badges/version/bulletsamplr)](https://cran.r-project.org/package=bulletsamplr)
 [![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
 [![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)](/commits/master)
 [![Travis build status](https://travis-ci.org/srvanderplas/bulletsamplr.svg?branch=master)](https://travis-ci.org/srvanderplas/bulletsamplr)
  [![Coverage status](https://codecov.io/gh/srvanderplas/bulletsamplr/branch/master/graph/badge.svg)](https://codecov.io/github/srvanderplas/bulletsamplr?branch=master)
  
The goal of bulletsamplr is to create bullet signatures which can be used to examine the behavior of match statistics under truly random conditions.

## Installation

You can install the released version of bulletsamplr from [github](https://github.com/srvanderplas/bulletsamplr) with:

``` r
devtools::install_github("srvanderplas/bulletsamplr")
```

## Example

The threshold bootstrap creates new sequences by resampling "cycles" from actual data. These cycles are created as follows:
```{r, fig.cap = "Bullet Signature"}
library(ggplot2)
library(bulletsamplr)
library(dplyr)

data(sig)
ggplot(aes(x = x, y = sig), data = sig) + 
  geom_line() + 
  ggtitle("Original Signature")
```

```{r, fig.cap = "Bullet Signature. Cycles are created when the sequence crosses the median, and include one positive region and one negative region. This ensures continuity between adjacent segments."}
threshold <- median(sig$sig, na.rm = T)
ggplot(aes(x = x, y = sig), data = sig) + 
  geom_line() + 
  geom_hline(aes(yintercept = threshold), color = "red") + 
  ggtitle("Threshold used to create cycles for future sample signatures")
```


```{r, fig.cap = ""}
sig_slices <- crosscut_slice(sig) %>% bind_rows() %>%
  group_by(chunk) %>%
  mutate(x = x - mean(x, na.rm = T))
sig_slices$chunk_label <- factor(sig_slices$chunk, labels = c("boundary_start", paste("cycle", unique(sig_slices$chunk)[-c(1, max(sig_slices$chunk))]), "boundary_end"))

ggplot(aes(x = x, y = sig), data = sig_slices) + 
  facet_wrap(~chunk_label, scales = "free") + 
  geom_line()
```
