---
title: "Exploring Sequence Matching with No Overlapping Regions"
author: "Susan VanderPlas"
date: "10/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(odbc)
library(bulletsamplr)
library(bulletxtrctr)
library(ggplot2)
library(stringr)
library(tidyr)
library(dbplyr)
library(randomForest)
```
```{r sim-setup}
con <- dbConnect(odbc::odbc(), "bullets", timeout = 10)
N <- 300
nshow <- 6
```

One way around the matching set problem is to divide the barrels up into two groups, and compare sequences assembled from each group. 

```{r}
barrels <- tbl(con, "bullet.slice") %>%
  collect() %>%
  mutate(barrel = str_replace_all(id, ".*_(\\d{2})-.*", "\\1"))

evens <- filter(barrels, barrel %in% c("02", "04", "06", "08", "10"))
odds <- filter(barrels, barrel %in% c("01", "03", "05", "07", "09"))

set.seed(2090490723)
evenseq <- purrr::map_df(1:N, ~crosscut_assemble(2000, evens), .id = "seq") %>%
  mutate(seqtype = 'even')
oddseq <- purrr::map_df(1:N, ~crosscut_assemble(2000, odds), .id = "seq") %>%
  mutate(seqtype = 'odd')
```

Plotting the first `r nshow` sequences from each set, 

```{r, fig.cap = 'First few sequences assembled. Even sequences were assembled from even-numbered barrels in Hamby sets; odd sequences from odd numbered barrels. This guarantees that there is no possibility that two sequences which are compared have chunks drawn from the same region of the same barrel.'}
first5 <- bind_rows(
  filter(evenseq, seq %in% as.character(1:nshow)),
  filter(oddseq, seq %in% as.character(1:nshow))
)

first5 %>%
  ggplot(aes(x = x, y = sig, color = seqtype)) + geom_line() + 
  facet_wrap(~seq)
```

We can now align each sequence:

```{r, fig.cap = 'First few sequences after alignment'}

aligned <- purrr::map2(tidyr::nest(evenseq, -seqtype, -seq)$data,
                       tidyr::nest(oddseq, -seqtype, -seq)$data,
                       ~ sig_align(.x$sig, .y$sig))

aligned_df <- data_frame(
  ccf = sapply(aligned, function(x) x$ccf),
  lag = sapply(aligned, function(x) x$lag),
  df = purrr::map(aligned, ~.$lands)
)

first5 <- aligned_df[1:nshow,] %>%
  mutate(i = 1:n()) %>%
  unnest() %>%
  gather(key = seqtype, value = sig, sig1:sig2)
  
first5 %>%
  ggplot(aes(x = x, y = sig, color = seqtype)) + geom_line() + 
  facet_wrap(~i)
```


We can also examine the maximum ccf of each aligned sequence:
```{r, fig.cap = 'Distribution of maximum cross-correlation for randomly assembled sequences with no possibility of same-barrel overlap'}
aligned_df %>%
  ggplot(aes(x = ccf)) + geom_density() + 
  geom_rug()
```

We can see that the ccf of each aligned sequence varies between .1 and .6, with the majority of sequences having a ccf of .3 - .4. 

## Null Feature Distribution

Using the aligned sequences, we can calculate striae alignment and features:
```{r}
striae <- purrr::map(aligned, sig_cms_max, span = 75)

striae_df <- data_frame(
  maxCMS = sapply(striae, function(x) x$maxCMS),
  ccf = sapply(striae, function(x) x$ccf),
  lag = sapply(striae, function(x) x$lag),
  lines = purrr::map(striae, ~.$lines),
  lands = purrr::map(striae, ~.$lands)
)

features <- purrr::map_df(striae, extract_features_all_legacy, resolution = 0.625)
features$rfscore <- predict(newdata = features, object = rtrees, type = 'prob')[,2]
```

The null distribution of random forest score is:
```{r}
ggplot(data = features, aes(x = rfscore)) + 
  geom_histogram(color = "black") + 
  scale_x_continuous("Random Forest Match Score") + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Random Forest Match Score")
```

### Cross-Correlation

```{r}
ggplot(data = features, aes(x = ccf)) + 
  geom_histogram(color = "black") + 
  scale_x_continuous("Cross-Correlation") + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Cross Correlation")
```
### Rough Correlation

```{r}
ggplot(data = features, aes(x = rough_cor)) + 
  geom_histogram(color = "black") + 
  scale_x_continuous("Rough Correlation") + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Rough Correlation")
```
### Sequence Distance

```{r}
ggplot(data = features, aes(x = D)) + 
  geom_histogram(color = "black") + 
  scale_x_continuous("Sequence Distance") + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Sequence Distance")
```
### Standard Deviation of Sequence Distance

```{r}
ggplot(data = features, aes(x = sd_D)) + 
  geom_histogram(color = "black") + 
  scale_x_continuous("Standard Deviation of Sequence Distance") + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Standard Deviation of Sequence Distance")
```

### Matches
```{r}
ggplot(data = features, aes(x = matches)) + 
  geom_histogram(color = "black", binwidth = 1) + 
  scale_x_continuous("Matches", breaks = c(0:5)*2) + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Matches")
```

### Mismatches
```{r}
ggplot(data = features, aes(x = mismatches)) + 
  geom_histogram(color = "black", binwidth = 1) + 
  scale_x_continuous("Mismatches", breaks = c(0:10)*2) + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Mismatches")
```

### Consecutive Matching Striae

```{r}
features %>%
  ggplot(aes(x = cms)) + 
  geom_histogram(color = 'black', binwidth = 1) + 
  scale_x_continuous("Maximum number of Consecutive Matching Striae", breaks = c(0:5) * 2) + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Maximum Consecutive Matching Striae")
```

### Nonconsecutive Matching Striae


```{r}
features %>%
  ggplot(aes(x = non_cms)) + 
  geom_histogram(color = 'black', binwidth = 1) + 
  scale_x_continuous("Nonconsecutive Matching Striae", breaks = c(0:10) * 2) + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Nonconsecutive Matching Striae")
```

### Sum of Peaks


```{r}
features %>%
  ggplot(aes(x = sum_peaks)) + 
  geom_histogram(color = 'black') + 
  scale_x_continuous("Sum of Aligned Striae") + 
  ylab("Number of Sequences") + 
  ggtitle("Null Distribution of Sum of Aligned Striae")
```
